---
title: "Paoer1"
subtitle: "TBD"
author: "TBD"
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(here)
library(janitor)
library(kableExtra)


all_data_path <- 
  here::here('inputs/data/raw_data.csv')

```

```{r, include=FALSE}

all_data <- read_csv(all_data_path)


commDiseases <- 
  all_data %>%
  janitor::clean_names()%>%
  select(disease:x85) %>%
  mutate(
         y0_19 = x1_2 + x01_04 + x05_09 + x10_14 + x15_19,
         y20_39 = x20_24+ x25_29 + x30_34 + x35_39, 
         y40_59 = x40_44+ x45_49 + x50_54 + x55_59,
         y60_79 = x60_64+ x65_69 + x70_74 + x75_79,
         y80 = x80_84 + x85) %>%
  mutate(total=y0_19+y20_39+y40_59+y60_79+y80)%>%
  select (disease,sex,y0_19,y20_39,y40_59,y60_79,y80,total)


ByDisease <- filter(commDiseases, !is.na(sex))%>%
  group_by(disease)%>%
  summarise(
    totalDisease= sum(total, na.rm = TRUE)
    ) %>% #take out the explanations in the bottom
  arrange(desc(totalDisease)) %>%
  top_n(3) # Top 3 disease by the number of people infected are Chlamydia,Gonorrhea,Influenza. 

top_3_disase <- ByDisease$disease


```

# Introduction

# Data

Our data is of penguins (Figure \@ref(fig:bills)) and we analyzed using R [@citeR] and the 'ggplot2' package[@cy1]. We build on [@diagnostics10040243] and [@covid2020post] to analyze the data.

```{r, echo = FALSE}
filter(commDiseases, !is.na(sex))%>%
  select(!sex) %>%
  group_by(disease)%>%
  summarise_each(list(sum)) %>%
  filter(disease %in% c(top_3_disase)) %>% 
  rename('Age 0-19' = y0_19, 'Age 20-39' = y20_39,'Age 40-59' = y40_59,'Age 60-79' = y60_79, 'Age 80+' = y80, 'Total' = total, 'Disease' = disease) %>% 
  kableExtra::kbl(caption = "Count of entries by year for 2017-2019") %>%
  kableExtra::kable_styling()
```

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it.

# Model

```{=tex}
\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}
```
Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# Appendix {.unnumbered}

\newpage

# References
